Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2018-05-22 2018/Journal of Supercomputing/JoS_Paper.tex
Last version

\begin{abstract}
This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains. Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced to solving a set of linked one-dimensional problems in parallel. The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one. The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions. The efficiency of the parallel algorithm was confirmed experimentally by solving numerically of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem’s functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions). It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems. 
%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$ for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) (1) by solving in parallel the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor/node. The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.

@@

The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes). In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect the total computational costs predominately in this case, and the difference between the time speedup and the iteration speedup will not be so great.
%Описание инфраструктуры

@@

The parallel index method for solving constrained global optimization problems considered in this paper offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domains);
	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

@@

%\bibitem{Evtushenko2009} Evtushenko, Yu.G., Malkova, V.U., Stanevichyus, A.A.: Parallel global optimization of functions of several variables. Comput. Math. Math. Phys. 49(2), 246--260 (2009)

@@

%\bibitem{Sergeyev2006} Sergeyev, Ya.D., Kvasov, D.E.: Global search based on efficient diagonal partitions and a set of Lipschitz constants. SIAM J. Optim. 16(3), 910--937 (2006)

%\bibitem{Paulavicius2014} Paulavi\v cius, R., Sergeyev, Y.D., Kvasov, D.E., \v Zilinskas, J. Globally-biased \textsc{Disimpl} algorithm for expensive global optimization. J. Glob. Optim. 59(2), 545--567 (2014)

%\bibitem{Sergeyev2015} Sergeyev, Y.D., Kvasov, D.E.: A deterministic global optimization using smooth diagonal auxiliary functions. Communications in Nonlinear Science and Numerical Simulation. 21(1-3), 99–111 (2015)

@@

%\bibitem{Barkalov2016} Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

>>>>

\begin{abstract}
This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains. Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced to solving a set of linked one-dimensional problems in parallel. The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one. The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions. The efficiency of the parallel algorithm was confirmed experimentally by solving numerically several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem's functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions). It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems. 
%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by solving in parallel the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor/node. The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.

@@

The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes). In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.
%Описание инфраструктуры

@@

The parallel index method for solving constrained global optimization problems considered in this paper offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

@@

@@

@@
